---
title: "Collaborative Research Project - Assignment 3"
bibliography:
date: "24 October 2014"
output: pdf_document
number_sections: yes
highlight: zenburn
authors: Meilin Moellenkamp and Nicolas Rosemberg
toc: yes
---


To automate the bibtex packages file:
repmis::LoadandCite(pkgs, file = 'RpackageCitations.bi
Note: Use a file name that is different from your literature BibTeX file!
b')

```{r, echo=FALSE, include=FALSE}

### 1. Setting Working Directory & Loading the data

# error=FALSE
# Setting the Working Directory
setwd("/Users/Meilin/Desktop/Collaborative Social Data/Research-Project-Assignment-3")
# setwd("/Users/Nico/Documents/Hertie/Social science data analysis/CollaborativeResearchProject")


# Loading the merged file from the Working directory
Merged <- read.csv(file="MergedData")
```


```{r, echo=FALSE, include=FALSE}

### 2. Load Required Packages

# install.packages("RJOSONIO")  
library(RJSONIO)
# install.packages("WDI")  
library(WDI)
# install.packages("dplyr")  
library(dplyr) 
# install.packages("tidyr")  
library(tidyr)
# install.packages("httr")  
library(httr) 
# install.packages("dplyr")  
library(dplyr)
# install.packages("XML")  
library(XML)
#install.packages("plyr")
library(plyr)
# install.packages("Amelia")  
library(Amelia) 
#install.packages("XLConnect")
library(XLConnect)    
# install.packages("countrycode")
library("countrycode")
# install.packages ("ggplot2")
library(ggplot2)
# # install.packages ("magrittr") ???
library(magrittr)
# install.packages ("fmsb")
library(fmsb)
# install.packages ("car")
library(car)
# install.packages("stargazer")
library(stargazer)


### 3. Check if Variables are coded as numeric
str(Merged)
# Code variables as numeric
Merged$Incidence <- as.numeric(Merged$Incidence)
Merged$DPT <- as.numeric(Merged$DPT)
Merged$Measles <- as.numeric(Merged$Measles)

```

# 1. Descriptive Statistics

For the discriptive statistics tables, plots and histograms will be shown in order to understand the distribution of the variables. The first table shows the main descriptive statistics of all the variables in the cleaned dataset (number of observations, mean, standard deviation, and min and max values).

```{r, echo=FALSE}
stargazer(Merged, type = "text", title="Table 1: Descriptive statistics", digits=1, out="table1.txt")
```

The histogram of the dependent variable (Figure 1) shows that the incidence rates are not normally distributed but strongly skewed to the left and only few incidence rates are higher than 1. 


```{r, echo=FALSE}

### 1. Plotting the dependent variable

# Histogram of dependent variable
Histdep <- hist(Merged$Incidence, breaks= seq(0,5,by=0.1), xlab="HIV Incidence Rate", main="Figure 1: Incidence Rate")
axis(side=1, at=seq(0,5, 0.5), labels=seq(0,5,0.5))
axis(side=2, at=seq(0,800,100), labels=seq(0,800,100))

```

In most regions of the world HIV incidences decreased between the period of 2000 to 2015 (see Figure 2).

```{r, echo=FALSE, include=FALSE}
# Look at incidence over time - seee the general trend
IncT <- ggplot(aes(x = year, y = Incidence), data = Merged) + geom_point(position = "jitter") + theme_bw() + geom_smooth()
options(digits=0)
IncT <- IncT + ggtitle("Figure 2: Incidence Rate over Time") + theme(plot.title = element_text(lineheight=3, face="bold"))
ggsave("fig2.png", width=5, height=5, dpi=100, device = pdf)
```

![Figure 2: Incidence Rate over Time] (/fig2.png)

When plotting the incidence rates per country (Figure 3) the countries with high incidence rates can be identified and the ranges of the observations per country are shown.

``` {r, echo=FALSE, include=FALSE}
# Look at incidence rate per country
IncC <- ggplot2::ggplot(Merged, aes(Incidence, country)) + geom_point() + theme_bw()
IncC + ggtitle("Figure 3: Incidence Rate per Country") + theme(plot.title = element_text(lineheight=3, face="bold"))

ggsave("fig3.png", width=8, height=10, dpi=100)

```

![Figure 3: Incidence Rate per Country] (/fig3.png)

As our research question is investigating why MDG 6.A is not being reached by some countries, we are not only interested in the general HIV incidence rate but also in the decrease in the incidence rate from 2000 to 2015. As stated in the introduction Target 6.A of the MDGs specifies that countries should "have halted by 2015 and begun to reverse the spread of HIV/AIDS"" (United Nations (2014)). 

For this purpose, the dependent variable was lagged by one period and the difference between the lag and the current year was calculated. Further, a dummy variable was created assigning a value of zero for those observations where the incidence rate decreased compared to the previous year or stayed the same (countries reaching MDG 6.A) and a value of one was assigned to those observations where the incidence rate increased (countries not reaching MDG 6.A). Figure 4 shows the direction of the change in the incidence rate compared to the previous year by country.

```{r, echo=FALSE, include=FALSE}
# Look at difference in incidence rate per country between one year to the next
IncClag <- ggplot2::ggplot(Merged, aes(IncidenceDif, country)) + geom_point() + theme_bw()
IncClag + ggtitle("Figure 4: Change in Incidence Rate compared to Previous Years by Country") + theme(plot.title = element_text(lineheight=3, face="bold"))

ggsave("fig4.png", width=8, height=10, dpi=100)
```

![Figure 4: Incidence Rate per Country] (/fig4.png)

Scatterplots were used for each category of the Dahlgrehn model in order to see whether the variables are skewed or multicollinear (Figures 5, 6 & 7).

```{r, echo=FALSE}
### 2. Plotting the independent variables

# Scatterplot of variables for socio-economic, cultural and environmental conditions
scatterplotMatrix(~ Incidence + + GDP + GDPpc + Rural + CO2 + LifeExpect + Population, 
                  transform=TRUE, data=Merged)

# Scatterplot of variables for living and working conditions 
scatterplotMatrix(~ Incidence + Water + Sanitation + Primary + Unemploym + HCexpendpc,
                  transform=TRUE, data=Merged)

# Scatterplot of variables for individual lifestyle factors
scatterplotMatrix(~ Incidence + DPT + Measles, transform=TRUE, data=Merged)

```


####################################################################################
########################### INFERENTIAL STATISTICS #################################
####################################################################################

As can be seen from the Scatterplots in the descriptive statistics most of the variables are not normally distributed. Further, the variables all have different scales. Therefore, the independent variables were logged for enabling comparisons in the inferential statistics part.

```{r, echo=FALSE, include=FALSE}
# Logging the independent variables for better comparability
Merged$lGDP <- log(Merged$GDP)
Merged$lGDPpc <- log(Merged$GDPpc)
Merged$lRural <- log(Merged$Rural)
Merged$lCO2 <- log(Merged$CO2)
Merged$lHCexpend <- log(Merged$HCexpend)
Merged$lWater <- log(Merged$Water)
Merged$lSanitation <- log(Merged$Sanitation)
Merged$lUnemploym <- log(Merged$Unemploym)
Merged$lPrimary <- log(Merged$Primary)
Merged$lHCexpendpc <- log(Merged$HCexpendpc)
Merged$lFemUnempl <- log(Merged$FemUnempl)
Merged$lFemSchool <- log(Merged$FemSchool)
Merged$lLifeExpect <- log(Merged$LifeExpect)
Merged$lDPT <- log(Merged$DPT)
Merged$lMeasles <- log(Merged$Measles)
```

```{r, echo=FALSE, include=FALSE}
# Creating a variable with the inverse from FemUnempl
Merged$Inverse <- 1/Merged$FemUnempl
Merged$Inverse <- log(Merged$Inverse)
```

As the dependent variable is coded as a dummy, being 0 if MDG 6.A was fulfilled and 1 for countries that don't fulfill MDG 6.A logistic regressions are used for the inferential statistics part. As Odds and Odds ratios are difficult to present to an audience that is not only consisting of statistical experts predicted probabilities are calculated following the logistic regressions. The interpretation of the results will only focus on the predicted probabilities.

```{r, echo=FALSE, include=FALSE}
# MODEL 1 - Running a general logistic regression using all independent variables
L1 <- glm(DDif ~ lGDP + lGDPpc + lRural + lCO2 + lHCexpend + lWater + lSanitation + lUnemploym + lPrimary + lHCexpendpc + Inverse + lFemSchool + lLifeExpect + lDPT + lMeasles,
          data=Merged, family = 'binomial')
summary(L1)
```

```{r, echo=FALSE}
# Testing for multicollinearity
vif(L1)
```

The test for variance inflation factors showed that in our first logistic regression model six variables showed high multicollinearity and had a heigher variance inflation than the threshold of 10. We tested the multicollinearity between the variables and found that there was high multicollinearity between the GDP and GDP per capital, Unemployment and Female unemployment, Primary education and female schooling. Therefore, we excluded one of these multicollinear variables for each group based on their explanatory strength for our research question, namely unemployment, primary education and GDP.

We tested the variance inflation factors of the new logistic regression model and all remaining variables passed the test.

```{r, echo=FALSE, include=FALSE}
# MODEL 2 - GDP
# Running the regression without multicollinear variables
L2 <- glm(DDif ~ lGDP + lRural + lCO2 + lHCexpend + lWater + lSanitation + lFemUnempl + lFemSchool + lLifeExpect + lDPT + lMeasles,
          data=Merged, family = 'binomial')

summary(L2)
```

```{r, echo=FALSE} 
# Testing for multicollinearity
vif(L2)
```

```{r, echo=FALSE} 
# MODEL 3 - GDPpc
L3 <- glm(DDif ~ lGDPpc + lRural + lCO2 + lHCexpend + lWater + lSanitation + lFemUnempl + lFemSchool + lLifeExpect + lDPT + lMeasles,
          data=Merged, family = 'binomial')
summary(L3)
# Testing for multicollinearity
vif(L3)

# Looking at confidence intervals
confint(L3)

```


```{r, echo=FALSE, include=FALSE}
# MODEL 4 - Including a new variable with FemSchool by quintiles

# Creating a new variable with FemSchool by quintiles
Merged$QFemSchool <- Merged$lFemSchool
Merged$QFemSchool[Merged$lFemSchool<=4.53] <-1
Merged$QFemSchool[Merged$lFemSchool>4.53 & Merged$lFemSchool<=4.622] <-2
Merged$QFemSchool[Merged$lFemSchool>4.622 & Merged$lFemSchool<=4.697] <-3
Merged$QFemSchool[Merged$lFemSchool>4.697] <-4

# LOGISTIC REGRESSION - Regressing the model with FemSchool quintiles
L4 <- glm(DDif ~ lGDPpc + lRural + lCO2 + lHCexpend + lWater + lSanitation + Inverse + as.factor(QFemSchool) + lLifeExpect + lDPT + lMeasles,
            data=Merged, family = 'binomial')
summary(L4)

# Look at Residual vs. Fitted Plot
plot(L4)

# PREDICTED PROBABILITIES:
#Regressing the model on FemSchool fixing the other indendent variables at Uganda's mean

# Looking at Uganda's average values
Uganda <- subset(Merged,(country=="Uganda"))
summary(Uganda)
summary(Merged)

# Regressing the model on FemSchool fixing the other indendent variables at Uganda's mean
fitted_L4 <- with(Merged,
                  data.frame(lGDPpc = 7.003,
                             lRural = 4.461,
                             lCO2 = -2.52,
                             lHCexpend = 2.125,
                             lLifeExpect = 3.98,
                             lWater = 4.18,
                             lSanitation = 3.46,
                             lDPT = 4.199,
                             lMeasles = 4.235,
                             Inverse = -1.2576,
                             QFemSchool = factor(1:4)))
fitted_L4


# Calculating the predicted probabilities
fitted_L4$fitted_L4_final <- predict(L4, newdata = fitted_L4,
                           type = 'response')
fitted_L4_final

# MODEL 5 - with the new independent variable without the interaction

# Creating a new variable with Inverse by quintiles
Merged$QInverse <- Merged$Inverse
Merged$QInverse[Merged$Inverse<=-2.534] <-1
Merged$QInverse[Merged$Inverse>-2.534 & Merged$Inverse<=-2.092] <-2
Merged$QInverse[Merged$Inverse>-2.092 & Merged$Inverse<=-1.459] <-3
Merged$QInverse[Merged$Inverse>-1.459] <-4

# Regressing the model with Inverse quintiles
L5 <- glm(DDif ~ lGDPpc + lRural + lCO2 + lHCexpend + lWater + lSanitation + as.factor(QInverse) + lFemSchool + lLifeExpect + lDPT + lMeasles,
          data=Merged, family = 'binomial')
summary(L5)

# Regressing the model on QInverse fixing the other indendent variables at Uganda's mean
fitted_L5 <- with(Merged,
                  data.frame(lGDPpc = 7.003,
                             lRural = 4.461,
                             lCO2 = -2.52,
                             lHCexpend = 2.125,
                             lLifeExpect = 3.98,
                             lWater = 4.18,
                             lSanitation = 3.46,
                             lDPT = 4.199,
                             lMeasles = 4.235,
                             lFemSchool = 4.808,
                             QInverse = factor(1:4)))
fitted_L5

# Calculating the predicted probabilities
fitted_L5$fitted_L5_final <- predict(L5, newdata = fitted_L5,
                           type = 'response')
fitted_L5_final


# MODEL 6 - Regressing the model with an interaction of femUnemployment and femSchool

# Creating a new variable with the interaction between FemSchool and Inverse
Merged$Interaction <- Merged$lFemSchool * Merged$Inverse
summary(Merged)
as.numeric(Merged$Interaction)

# Creating a new variable with the interaction by quintiles
Merged$QInteraction <- Merged$Interaction
Merged$QInteraction[Merged$Interaction<=-11.469] <-1
Merged$QInteraction[Merged$Interaction>-11.469 & Merged$Interaction<=-9.508] <-2
Merged$QInteraction[Merged$Interaction>-9.508 & Merged$Interaction<=-6.82] <-3
Merged$QInteraction[Merged$Interaction>-6.82] <-4

# LOGISTIC REGRESSION:
L6 <- glm(DDif ~ lGDPpc + lRural + lCO2 + lHCexpend + lWater + lSanitation + lLifeExpect + lDPT + lMeasles + as.factor(QInteraction) + Inverse + lFemSchool,
          data=Merged, family = 'binomial')
summary(L6)


# Regressing the model on FemSchool fixing the other indendent variables at Uganda's mean
fitted_L6 <- with(Merged,
                  data.frame(lGDPpc = 7.003,
                             lRural = 4.461,
                             lCO2 = -2.52,
                             lHCexpend = 2.125,
                             lLifeExpect = 3.98,
                             lWater = 4.18,
                             lSanitation = 3.46,
                             lDPT = 4.199,
                             lMeasles = 4.235,
                             Inverse = -1.2576,
                             lFemSchool = 4.808,
                             QInteraction = factor(1:4)))
fitted_L6


# Calculating the predicted probabilities
fitted_L6$fitted_L6_final <- predict(L6, newdata = fitted_L6,
                            type = 'response')
fitted_L6_final


# plot(L5)

# Using the anova function to evaluate the effect on the deviance of adding the interaction term 
# anova(L3,L4)

# anova(L4,test="Chisq")
```

The results of the logistic regressions were exported with Stargazer. Table 2 shows the logistic regression results. The results table shows that out of all the tested variables in the model only the variable Sanitation has a significant impact on changes in HIV/Aids incidence rates fo.


```{r, echo=FALSE, type = 'latex', header = FALSE}
#### SAVING THE REGRESSION OUTPUTS WITH STARGAZER
# Regression Results of Logistic Regression
stargazer(L4, L5, L6, title="Table 2: Logistic Regression Results", type="text",align=TRUE, out="OddsResults.txt")
```


```{r, echo=FALSE, type = 'latex', header = FALSE}
# Predicted Probabilities Model 1
stargazer(fitted_L4_final, title="Predicted Probabilities Model 1", type="text",align=TRUE, out="models.txt")
```

```{r, echo=FALSE, type = 'latex', header = FALSE}
# Predicted Probabilities Model 2
stargazer(fitted_L4_final, title="Predicted Probabilities Model 2", type="text",align=TRUE, out="models.txt")
```

```{r, echo=FALSE, type = 'latex', header = FALSE }
# Predicted Probabilities Model 3
stargazer(fitted_L4_final, title="Predicted Probabilities Model 3", type="text",align=TRUE, out="models.txt")
```
If you want to automatically generate tables from regression model output objects, texreg is a good package to turn to. First estimate your models:

L1 <- glm(admit ~ gre,
              data = Admission, family = 'binomial')

L2 <- glm(admit ~ gre + gpa,
              data = Admission, family = 'binomial')

L3 <- glm(admit ~ gre + gpa + as.factor(rank),
              data = Admission, family = 'binomial')
Then use the stargazer function to create a results table. For PDFs set type = 'latex'. There are many stylistic modifications you can make with this function.

# Create cleaner covariate labels
labels <- c('(Intercept)', 'GRE Score', 'GPA Score', '2nd Ranked School', '3rd Ranked School',
            '4th Ranked School')

stargazer::stargazer(L1, L2, L3, covariate.labels = la title = 'Logistic Regression Estimates of Grad Sch digits = 2, type = 'latex', header = FALSE)



LIMITATIONS

The paper had to make some compromises regarding its original aims that were outlined in the first research proposal.

1. Due to huge amounts of missing data and some problems with multicollinearity lots of variables had to dropped and could ultimately not be integrated in the logistic regression models. Therefore, the categories that were supposed to be tested 


are facing some limitations 
A simpler and more common model on the main determinants of health is the ???rainbow model???, developed by Dahlgren and Whitehead (Dahlgren and Whitehead (1991), p.11). This model gives an overview of the main health determinants, reflecting the relationship between the individual, its environment and different health outcomes. Individuals are at the centre of the model with a set of fixed biological and genetical preconditions. Building upon these, four layers of influence on health can be identified: individual lifestyle factors, social and community networks, living and working conditions and general socio-economic, cultural and environmental conditions.